Holomorphic field theories and higher algebra

### 2022-08-12 draft.tex
section 1

\spacing{1.25}

@@

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\pAA{\mathring{\AA}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}

@@

@@

The mathematical motivation is simple:
for most classical holomorphic field theories, the solutions to the equations of motion form a moduli space of natural interest to complex geometers,
and it is natural to hope that its quantization is equally interesting.
A well-known example is holomorphic Chern-Simons theory,
which lives on a Calabi-Yau 3-fold $X$ and involves the {\em derived} moduli space $\RR\Bun_G(X)$ of holomorphic principal $G$-bundles on $X$.
As a somewhat mundane example, let $\RR\Hol(X,V)$ denote the derived moduli space of holomorphic maps from some complex manifold $X$ to another complex manifold $V$.
(For $V$ a vector space, a ``derived'' enhancement is familiar as the higher dimensional cohomology $H^*(X, \cO) \otimes V$.)
One can take a ``shifted cotangent bundle'' $T^*[-1] \RR\Hol(X,V)$,
and this derived moduli space encodes solutions to
a theory known as the $\beta\gamma$ system with target $V$.
(The name is silly: it's due to the original notation people used.)
Every question one asks about the classical physics of the $\beta\gamma$ theory translates into a question about this moduli space.
Other classes of examples are afforded by holomorphic variants of familiar AKSZ $\sigma$-models; the relevant moduli space being the derived mapping space between complex manifolds.

A physical motivation comes from supersymmetric field theories, 
which (loosely speaking) are theories on $\RR^n$ whose symmetries include not only the isometries of $\RR^n$, but an extension to a Lie supergroup known as a super Poincar\'e group.
%The relevant extension is often labeled by $\cN=1, 2,$ and so on, 
%with larger $\cN$ denoting a bigger supergroup. 
Thanks to the enhanced symmetry, such theories can often be understood in greater detail than non-supersymmetric theories.
The relevance to the models we study is that supersymmetric theories admit ``twists'' (in essence, deformations) that are holomorphic field theories, 
even when they do not admit twists to TFTs (such as the famed A- and B-models of mirror symmetry) \cite{CosHol}.
As an example, four-dimensional minimally supersymmetric Yang-Mills theory admits no topological twist, but does admit a twist to holomorphic field theory whose moduli space is described, in part, by $\RR \Bun_G(X)$. 
It is natural to wonder how the myriad tools and computations deployed on the supersymmetric theory carry over to this twist and conversely whether information can flow back from a holomorphically twisted theory to the untwisted one, 
much as TFTs have enriched the understanding of supersymmetric theories.
In short, a holomorphic twist contains the ``holomorphic sector'' of the untwisted theory just as a topological twist contains its ``topological sector.''

As a final motivational remark, we note in one complex dimension (i.e., working on Riemann surfaces), holomorphic field theories appear as chiral conformal field theories (CFTs) on 2-dimensional manifolds --- whose manifestations in mathematics include vertex algebras, affine Lie algebras, and loop groups --- and have had a strong impact in mathematics.
It is natural to explore higher-dimensional analogs.

@@

%complex geometry, dolbeault, dg stuff.

@@

Let us begin with the example of holomorphic Chern--Simons theory, 
which we have already said describes the moduli space of Maurer--Cartan elements describe deformations of the trivial holomorphic $G$-bundle. 
When $X$ is a Calabi--Yau three-fold and $\fg$ is equipped with an invariant pairing, 
this formal moduli space has a very special property: 
it arises from the variational problem of an action functional, namely
\[
S = \int_X \Omega \wedge {\rm CS}(A) 
\]
where $\Omega$ is the holomorphic volume form and ${\rm CS}(A)$ is the Chern--Simons Lagrangian for the connection $(0,1)$-form $A$.

@@

\owen{Mention main classes of field theories (scalar and sigma models, fermions as matter, gauge theories for ``forces'', gravity, and their combinations) and how there are holomorphic analogs}

@@

% beta gamma systems

\subsubsection{Holomorphic fermionic field theories}

% bc systems

@@

% BF and CS

@@

% curved beta gamma

@@

% Kodaira-Spencer

\subsection{Examples from supersymmetric field theories}

@@

\owen{Revise to explain SYM as source for some examples above, but also add brief discussions of newer examples like twisted SUGRA and exotic super Lie algebras}

@@

By definition, a supersymmetric field theory on $\RR^d$ is a theory that is acted upon by a super Poincar\'{e} algebra. 
We focus on Euclidean field theories and work in Riemannian signature,
hence for us the ``super Poincar\'{e} algebra'' is a super Lie algebra of the form
$\mathfrak{so}(d) \ltimes \ft$
where $\ft$ is the super Lie algebra of {\em supertranslations} whose even part $\ft^0 = \RR^d$ is the Lie algebra of ordinary translations and whose odd part $\ft^1$ is a sum of spin representations. 
While the Lie algebra $\RR^d$ of ordinary translations is abelian, the Lie algebra $\ft$ carries a nontrivial (super) Lie bracket,
and this Lie bracket is defined in terms of a $\mathfrak{so}(d)$-equivariant non-degenerate symmetric pairing
\[
\Gamma : {\rm Sym}^2(\ft^{1}) \to {\rm Sym}(\ft^0) \cong \RR^d 
\]
by the formula $[Q, Q'] = \Gamma(Q, Q')$.

@@

Building upon the physical literature, Elliott, Safronov, and the second author have given a complete characterization of the all twisted supersymmetric Yang--Mills theories in dimensions $2 \leq d \leq 10$ \autocite{ESW}. 
These twisted theories are analogs of BF theory and Chern--Simons theory, so long as one includes purely topological and purely holomorphic theories. 
We will describe a few examples in detail in the BV formalism in Section~\ref{sec:twistedsusybv},
but we want to indicate now what our main theorem implies about some interesting examples.

For our purposes here, a crucial property of a twist $Q$ is the dimension $k_Q$ of its image ${\rm Im} \; \Gamma(Q, \cdot) \subset \RR^d$,
which we call the number of {\em invariant} directions of~$Q$.
By the non-degeneracy of the pairing $\Gamma$, it follows that $k_Q \geq \frac{d}{2}$.
If the number of invariant directions is maximal with $k_Q = d$, then the twisted theory $\cT^Q$ is purely topological. 
If the number of invariant directions is minimal with $k_Q = \frac{d}{2}$, then the twisted theory $\cT^Q$ is purely holomorphic.

As a representative example of how twisting behaves, 
consider four-dimensional supersymmetric Yang--Mills theory with $\cN=2$ supersymmetry,
which means the odd part $\ft^1$ is eight-dimensional. 
%The sum of two copies of the spin representations.
%\owen{Did I get that right?}
We only consider here the pure gauge theory with Lie algebra~$\fg$. 
Then we have the bounds $2 \leq k_Q \leq 4$, and each value gives a different, well-known field theory:
\begin{itemize}
\item When $k_Q = 4$, the twisted theory is the topological field theory on $\RR^4$ known as {\it Donaldson--Witten theory} \autocite{WittenTQFT};
\item When $k_Q = 3$, the twisted theory is the THFT studied by Kapustin in \autocite{Kapustin};
\item When $k_Q = 2$, the twisted theory is holomorphic BF theory on $\CC^2$ with values in a graded Lie algebra $\fg[\ep]$, where $\ep$ is a parameter of cohomological degree~$1$.
\end{itemize}
We will describe these theories completely later in Section~\ref{sec:twistedsusybv},
but our main theorem implies the following.

@@

\owen{We should introduce the idea here and revisit later}

>>>>

\spacing{1.1}

@@

\def\g{{\mathfrak g}}
\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\pAA{\mathring{\AA}}
\def\jou{{\mathtt{j}}}
\def\bgamma{{\mathbf{\gamma}}}
\def\bbeta{{\mathbf{\beta}}}
\def\CDO{{\cC\cD}}
\def\SU{{\rm SU}}
\def\sfa{\mathsf{a}}
\def\del{\partial}
\def\til{\Tilde}

@@

\brian{we need to make sure to cite nekrasov}

@@

The mathematical motivation is simple:
for most classical holomorphic field theories, the solutions to the equations of motion form a moduli space of natural interest to complex geometers,
and it is natural to hope that its quantization is equally interesting.
A well-known example is holomorphic Chern-Simons theory,
which lives on a Calabi-Yau 3-fold~$X$ and involves the {\em derived} moduli space $\RR\Bun_G(X)$ of holomorphic principal $G$-bundles on~$X$.
As a somewhat mundane example, let $\RR\Hol(X,Y)$ denote the derived moduli space of holomorphic maps from some complex manifold $X$ to another complex manifold $Y$.
(For $Y = V$ a vector space, a ``derived'' enhancement is familiar as the higher dimensional cohomology $H^*(X, \cO) \otimes V$.)
One can take a ``shifted cotangent bundle'' $T^*[-1] \RR\Hol(X,V)$,
and this derived moduli space encodes solutions to
a theory known as the $\beta\gamma$ system with target $V$.
(The name is silly: it's due to the original notation people used.)
Every question one asks about the classical physics of the $\beta\gamma$ theory translates into a question about this moduli space.
Other classes of examples are afforded by holomorphic variants of familiar AKSZ $\sigma$-models; the relevant moduli space being the derived mapping space between complex manifolds.

A physical motivation comes from supersymmetric field theories, 
which (loosely speaking) are theories on $\RR^n$ whose symmetries include not only the isometries of $\RR^n$, but an extension to a Lie supergroup known as a super Poincar\'e group.
%The relevant extension is often labeled by $\cN=1, 2,$ and so on, 
%with larger $\cN$ denoting a bigger supergroup. 
Thanks to the enhanced symmetry, such theories can often be understood in greater detail than non-supersymmetric theories.
The relevance to the models we study is that supersymmetric theories admit ``twists'' (in essence, deformations) that are holomorphic field theories, 
even when they do not admit twists to TFTs (such as the famed A- and B-models of mirror symmetry) \cite{CosHol}.
As an example, four-dimensional minimally supersymmetric Yang-Mills theory admits no topological twist, but does admit a twist to holomorphic field theory whose moduli space is described, in part, by the moduli space of holomorphic $G$-bundles~$\RR \Bun_G(X)$. 
It is natural to wonder how the myriad tools and computations deployed on the supersymmetric theory carry over to this twist and conversely whether information can flow back from a holomorphically twisted theory to the untwisted one, 
much as TFTs have enriched the understanding of supersymmetric theories.
In short, a holomorphic twist contains the ``holomorphic sector'' of the untwisted theory just as a topological twist contains its ``topological sector.''

As a final motivational remark, we note in one complex dimension (i.e., working on Riemann surfaces), holomorphic field theories appear as chiral conformal field theories (CFTs) on two-dimensional manifolds --- whose manifestations in mathematics include vertex algebras, affine Lie algebras, and loop groups --- and have had a strong impact in mathematics.
It is natural to explore higher-dimensional analogs.

@@

A theory is usually described, in part, in terms of a space of fields which comprise some possibly nonlinear space like a mapping space or the space of connections. 
In perturbative field theory, which describes the formal neighborhood of the theory at points in such spaces, the fields $\Phi$ are typically described by the smooth sections of a bundle on spacetime.
The equations of motion which dictate the dynamics of the particular model are encoded by a partial differential equation heuristically of the form 
\[
\sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\]
where $D_k$ is a $k$-ary polydifferential operator.
As an example of a Euclidean theory, the $\phi^4$-theory on a Riemannian four-manifold $M$ is described by the equation
\[
\Delta \phi + \frac{1}{3!} \phi^3 = 0 
\]
where $\Phi = \phi$ is simply a smooth function on $M$.

In a {\em holomorphic theory} one starts with a complex manifold $X$ as the spacetime.
This means, in part, that the complexified tangent bundle of $X$ decomposes as a sum of vector bundles $\T_X \oplus \Bar{\T}_X$ where $\T_X$ (respectively $\Bar{T}_X$) is the holomorphic (respectively anti-holomorphic) tangent bundle.
In local coordinates sections of $\T_X$ have the form $f^i \del / \del z_i$ whereas sections of $\Bar{\T}_X$ have the form $f^i \del / \del \zbar_i$.

Globally, holomorphic theories are described like in the usual case by mapping spaces or spaces of connections.
The key feature in a holomorphic theory is that the equations of motion dictate that fields be holomorphic.
Perturbatively, where fields are sections of some holomorphic vector bundle $V$, we can make this precise. 
Prescribing a holomorphic structure on a vector bundle $V$ is equivalent to the data of a $\dbar$-operator
\beqn
\dbar \colon \Omega^{0,q}(X, V) \to \Omega^{0,q+1}(X,V) 
\eeqn
where $\Omega^{0,q}(X,V)$ is the space of smooth sections of the vector bundle $\wedge^q\Bar{T}_X^* \otimes V$. 
Holomorphicity of $V$ boils down to the requirement that $\dbar \circ \dbar = 0$.
The Dolbeault complex of $V$ is the resulting complex $(\Omega^{0,\bu}(X, V), \dbar)$ where $\Omega^{0,q}(X,V)$ sits in degree $+q$.
A standard result in complex geometry is that the Dolbeault complex is a resolution for the sheaf of holomorphic sections of the holomorphic vector bundle $V$.
In the case that $V$ is the trivial bundle $\ul{\CC}$ the Dolbeault complex $\Omega^{0,\bu}(X)$ is a resolution for the sheaf of holomorphic functions on $X$.

The fields of a holomorphic theory are given by elements of the Dolbeault complex $\Omega^{0,\bu}(X,V)$.
The equations of motion in a holomorphic theory take the form
\beqn
\dbar \Phi + \sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\eeqn
where $D_k$ are $k$-ary {\em holomorphic} polydifferential operators.
This means that in local coordinates $D_k$ only depend on the holomorphic coordinates $\{z_i\}$ and involve only holomorphic derivatives $\del / \del z_i$.

@@

Let us begin with the example of holomorphic Chern--Simons theory, 
which we have already said describes the moduli space of Maurer--Cartan elements describe deformations of the trivial holomorphic $G$-bundle. 
When $X$ is a Calabi--Yau three-fold and $\fg$ is equipped with an invariant pairing $\<\cdot , \cdot \>$, 
this formal moduli space has a very special property: 
it arises from the variational problem of an action functional, namely
\beqn
S = \int_X \Omega \wedge {\rm CS}(A) 
\eeqn
where $\Omega$ is the holomorphic volume form and 
\beqn
{\rm CS}(A) = \frac12 \< A ,\d A \> + \frac16 \<A , [A,A]\>
\eeqn
is the Chern--Simons Lagrangian for the connection $(0,1)$-form $A$.
Notice that due to the presence of the holomorphic volume form $\Omega$ on the $\dbar$ operator contributes to the kinetic term and so the action can be written as 
\beqn
S = \int_X \Omega \wedge \left(\frac12 \< A ,\dbar A \> + \frac16 \<A , [A,A]\>\right) .
\eeqn

@@

@@

The most elementary holomorphic field theory is a generalization of the so-called `$\beta\gamma$ system' which is primarily studied in the context of string theory.
The input data is simply a complex manifold $X$ of arbitrary complex dimension $n$ together with a (super) holomorphic vector bundle $V$ on $X$.
The fields of the theory consist of pairs 
\begin{align*}
\gamma & \in \Gamma(X,V)\\
\beta & \in \Omega^{n,n-1}(X,V^*) .
\end{align*}
The action functional is 
\beqn
\int_X \beta \wedge \dbar \gamma 
\eeqn
which utilizes the natural pairing between $V$ and $V^*$. 
The theory is free, just like the ordinary $\beta\gamma$ system on Riemann surfaces.

Notice that there is a linear symmetry present in the above action given by shifting $\beta$ by a $\dbar$-exact form
\beqn
\beta \mapsto \beta + \dbar \eta , \quad \eta \in \Omega^{n,n-2}(X,V^*)
\eeqn
For this model to be well-posed at the quantum level we must include the ghosts $\eta$.
Further, $\eta$ can be shifted by an exact form in a way that does not effect the above gauge transformation, so we must also include ghosts for ghosts, etc..

@@

We have already met holomorphic Chern--Simons theory, which is defined on Calabi--Yau three-folds.
There is a related holomorphic gauge theory which is defined in arbitrary complex dimensions and does not require the data of a Calabi--Yau structure.

If $n$ is the complex dimension of $X$, the fields consist of pairs $(A,B)$ where $A$ is a $(0,1)$ connection on a principal $G$-bundle $P \to X$ and
\beqn
B \in \Omega^{n,n-2}(X, \fg_P^*) ,
\eeqn
where $\fg_P$ is the adjoint bundle of $P$. 
The action functional is 
\beqn
S(A,B) = \int_X B \wedge F_A 
\eeqn
where $F_A$ is the curvature of the connection $A$.
From this formula, it is clear that the model we are describing is a holomorphic analog of topological $BF$ theory. 
Notice that by form types this action can be written as
\beqn
S(A,B) = \int_X B \wedge \left(\dbar A + \frac12 [A,A] \right) .
\eeqn

@@

If $X,Y$ are any two complex manifolds there is a theory whose fields consist, in part, of the derived moduli space of smooth maps $\RR {\rm Map}(X,Y)$ and whose equations of motion dictate that such maps be holomorphic.
This is a nonlinear generalization of the $\beta\gamma$ system we just introduced.

The precise space of fields, in cohomological degree zero, consist of pairs $(\gamma, \beta)$ where $\gamma \colon X \to Y$ is a smooth map and 
\[
\beta \in \Omega^{n,n-1}(X, f^* \T^*_Y) 
\]
where $\T^*_Y$ is the holomorphic cotangent bundle of $Y$. 
The action functional is the familiar one
\beqn
\int_X \beta \dbar \gamma .
\eeqn

@@

Perturbatively, any model of gravity describes, in part, deformations of the moduli space of metric structures on a given manifold. 
There is a holomorphic avatar of this which describes deformations of complex structure.
The fields consist of pairs $(\mu, \eta)$ where
\begin{align*}
\mu & \in \Omega^{0,1}(X, \T_X) \\
\nu & \in \Omega^{0,n-2}(X, K_X \otimes \T^*_X) .
\end{align*}
The action functional is
\beqn
\int_X \left(\eta \dbar \mu + \frac12 \eta [\mu,\mu]\right)
\eeqn
where $[\mu, \mu] \in \Omega^{0,2} (X, \T_X)$ stands for the Lie bracket of holomorphic vector fields extended to the Dolbeault valued sections of $\T_X$ in the natural way.

In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure where we denote by $\Omega \in \Omega^{3,hol}(X)$ the holomorphic volume form.
Kodaira--Spencer theory is the holomorphic theory whose fields consist of $(0,1)$ forms on $X$ with values in the holomorphic tangent bundle $\T_X$
\beqn
\mu \in \Omega^{0,1}(X,\T_X) .
\eeqn 
A crucial assumption of this model is that we must assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning its divergence vanishes ${\rm div}_\Omega \mu = 0$.
The action functional is very similar to the Chern--Simons action
\beqn
S(\mu) = \int_X \left(\frac12 \mu \dbar \del^{-1} \mu + \frac13 \mu^3\right) .
\eeqn
The integrand is not defined globally.
In the integral we use the Calabi--Yau form to identify the integrand with 
\brian{finish}

\subsection{Local version and examples from supersymmetric field theories}

@@

There is the following local characterization of a holomorphic field theory which allows us to connect directly to supersymmetric field theories. 
Suppose that our spacetime is $\CC^n = \RR^{2n}$ and assume that the theory we are considering is translation invariant, meaning that there is an action buy the $2n$-dimensional Lie algebra spanned by the vector fields $\del / \del {z_i}, \del / \del {\zbar_i}$ for $i=1,\ldots,n$. 
In a holomorphic theory we require that the anti-holomorphic vector fields $\del / \del {\zbar_i}$ act homotopically trivial. 
In the derived world this is prescribed by the data of homotopy trivializations for such vector fields.

@@

By definition, a supersymmetric field theory on $\RR^d$ is a theory that is acted upon by a super Poincar\'{e} algebra. 
We focus on Euclidean field theories and work in Riemannian signature,
hence for us the ``super Poincar\'{e} algebra'' is a super Lie algebra of the form
$\mathfrak{so}(d) \ltimes \ft$
where $\ft$ is the super Lie algebra of {\em supertranslations} whose even part $\ft^0 = \RR^d$ is the Lie algebra of ordinary translations and whose odd part $\ft^1$ is a sum of spin representations. 
While the Lie algebra $\RR^d$ of ordinary translations is abelian, the Lie algebra $\ft$ carries a nontrivial (super) Lie bracket,
and this Lie bracket is defined in terms of a $\mathfrak{so}(d)$-equivariant non-degenerate symmetric pairing
\beqn
\label{e:Gamma}
\Gamma \colon {\rm Sym}^2(\ft^{1}) \to {\rm Sym}(\ft^0) \cong \RR^d 
\eeqn
by the formula $[Q, Q'] = \Gamma(Q, Q')$.

@@

Building upon the physical literature, Elliott, Safronov, and the second author have given a complete characterization of the all twisted supersymmetric Yang--Mills theories in dimensions $2 \leq d \leq 10$ \autocite{ESW}. 
These twisted theories are analogs of BF theory and Chern--Simons theory, so long as one includes purely topological and purely holomorphic theories.
In even dimensions $d = 2n$, basic properties of the supersymmetry algebra guarantee that there exists a twist rendering the theory holomorphic.\footnote{This fails in one example where the dimension and amount of supersymmetry is too small, namely for $\cN=(1,0)$ supersymmetry in two dimensions.}
Indeed, so long as a twisted supercharge is nonzero $Q \ne 0$ one see that the image of the map $[Q,-] = \Gamma(Q,-)$ from \eqref{e:Gamma} is at least $n$-dimensional.
A {\em minimal} or {\em holomorphic} supercharge is one for which the dimension is exactly $n$ and we declare the the image of $Q$ is spanned by the anti-holomorphic vector fields $\{\del / \del \zbar_i\}$.
This fits with the local characterization mentioned at the beginning of the subsection since in this case there exists homotopies $\{\eta_i\}$ with the property that 
\beqn
[Q,\eta_i] = \frac{\del}{\del \zbar_i} .
\eeqn

%For our purposes here, a crucial property of a twist $Q$ is the dimension $k_Q$ of its image ${\rm Im} \; \Gamma(Q, \cdot) \subset \RR^d$,
%which we call the number of {\em invariant} directions of~$Q$.
%By the non-degeneracy of the pairing $\Gamma$, it follows that $k_Q \geq \frac{d}{2}$.
%If the number of invariant directions is maximal with $k_Q = d$, then the twisted theory $\cT^Q$ is purely topological. 
%If the number of invariant directions is minimal with $k_Q = \frac{d}{2}$, then the twisted theory $\cT^Q$ is purely holomorphic.

As a representative example of how twisting behaves, 
consider pure four-dimensional supersymmetric Yang--Mills theory with $\cN=1$ supersymmetry.
The odd part of the supersymmetry algebra $\ft^1$ is four-dimensional and there exists a unique holomorphic twist $Q$ up to equivalence. 
%The sum of two copies of the spin representations.
%\owen{Did I get that right?}

Then, the holomorphic twist

@@

\owen{We should introduce the idea here and revisit later}
\brian{Explain dimensional reduction vs compactification.
From topological AKSZ theory to holomorphic one.}
